Real-Time Scheduling Algorithms and Battery Consumption of Mobile Devices

### 2017-05-01 paper-RMDON.tex
un poco mas de secciones

\section{Introduction}
\IEEEPARstart{N}{owadays} real-time applications are taking over mobile devices more than ever and, although this brings a lot of new opportunities, it brings challenges too. There is a plenty of examples of applications that includes collective, social, and distributed features out there and, it just a matter to take a look to a smartphone of an average user to see several of those in action. But, despite the success behind the cell phones as mobile devices, currently there are many other hardware architectures which has mobile capabilities as well and, in which the implementation of new and interactive applications have increased in recent years. Weareable technolgy and ARM architectures are good examples of this.

\section{Real-Time Scheduling Algorithms}

\subsection{On mobile devices}

\section{Scheduling algorithm 1}
\section{Scheduling algorithm 2}
\section{Scheduling algorithm 3}

@@

@@

@@

@@

\bibitem{quian-andresen} H.~Qian and D.~Andresen, \emph{An energy-saving task scheduler for mobile devices}. 2015 IEEE/ACIS 14th International Conference on Computer and Information Science (ICIS), Las Vegas, NV, 2015, pp. 423-430.
doi: 10.1109/ICIS.2015.7166631

\bibitem{rao-et-al} V.~Rao, N.~Navet and G.~Singhal, \emph{Battery aware dynamic scheduling for periodic task graphs}. 2006 Proceedings 20th IEEE International Parallel \& Distributed Processing Symposium, Rhodes Island, 2006. doi: 10.1109/IPDPS.2006.1639403

\bibitem{takada-sakamura} H.~Takada and K.~Sakamura, \emph{Real-time synchronization protocols with abortable critical sections}. 1994 Proceedings of 1st International Workshop on Real-time Computing Systems \& Application, pp. 48-52

@@

\bibitem{wilke-et-al} C.~Wilke, S.~Richly, S.~Götz, C.~Piechnick and U.~Aßmann, \emph{Energy Consumption and Efficiency in Mobile Applications: A User Feedback Study}, 2013 IEEE International Conference on Green Computing and Communications and IEEE Internet of Things and IEEE Cyber, Physical and Social Computing, Beijing, 2013, pp. 134-141.
doi: 10.1109/GreenCom-iThings-CPSCom.2013.45

>>>>

\section{Introduction}
\IEEEPARstart{N}{owadays} real-time applications are taking over mobile devices more than ever and, although this brings a lot of new opportunities, it brings challenges too. There is a plenty of examples of applications that includes collective, social, distributed and even virtual/augmented reality features out there and, it just a matter to take a look to a smartphone of an average user to see several of them in action. But, besides the success behind laptops and cell phones as the mobile devices, currently there are many other hardware architectures which has mobile capabilities as well and, in which the implementation of new and interactive applications have increased in recent years. Weareable technolgy and ARM architectures are good examples of this.

Thanks to hardware advances, both companies and users are pushing to provide and experience better applications on mobile devices but this still has big impact in its performance, particularly in battery conpsumption. Research groups have focused on optimizing the hardware of mobile devices, as well as their middleware increasing both the devices' uptime and the user satisfaction but the greater demand in performance for both mobile device hardware and its applications conflicts with the desire for longer battery life.

Users are craving for longer battery life, some even claim that the next big thing in technology ought be better batteries\cite{moren}. Taking as example mobile applications, a study conducted by \cite{wilke-et-al} analyzed 9 millions user comments and 27,000 mobile applications under the Google Play store and demostrated that users of mobile applications are interested in enery-efficient applications and energy-inefficient applications lead to frustrated users to negative feedback and lower application ratings. Some other relevants results of the study revealed that: more than 18\% of the analyzed applications have user feedback comprising comments on energy consumption problems, energy-efficiency issues negatively influence the grades users give for mobile apps in market places and that free apps do not have more energy consumption problems than payed apps.

The energy use of a typical laptop computer, smartphone and a tablet is dominated by the backlight and display, disks and CPU. They use a number of techniques to reduce the energy consumed by the disk and display, primarily by turning them off after a period of no use. Power consumed by the CPU is becoming more significant because of the type of the applications they run now: data-intensive, multimedia, games and collaborative applications that need to be in a constant sync with networks, sensors, and other devices. Dynamic voltage scaling (DVS) is a common mechanism to save CPU energy. It exploits an important characteristic or CMOS\footnote{CMOS: Complementary metal–oxide–semiconductor}-based processors: the maximun frequency scales almost linearly to the voltage, and the energy consumed per cycle is proportional to the square of the voltage. A lower frequency hence enables a lower voltage and yields a quadratic energy reduction\cite{chandrakasan}. The major goal of DVS is to reduce energy by as much as possible without degrading appplication performance. The efectiveness of DVS techniques is, therefore, dependent on the ability to predict application CPU demands -- overestimating them can waste CPU and energy resources, while understimating them can degrade aplication performance\cite{yuan-nahrstedt}.

A real-time operating system has a well-specified maximum time for each action that it performs to support applications with precise timing needs. Systems that can guarantee these maximum times are called hard real-time systems. Those that meet these times most of the time are called soft real-time systems. Deploying an airbag in response to a sensor being actuated is a case where you would want a hard real-time system. Decoding video frames is an example of where a soft real-time system will suffice

In this paper, we present [a set of] [the following] real-time scheduling algorithms  [A, B, C] that have been proposed to improve energy-saving for mobile devices. We discuss their main features

\section{Real-Time Task Scheduling Algorithms}
Several schemes of classification of real-time task scheduling algorithms exist. A scheme in \cite{mall} classifies the real-time task scheduling algorithms based on how the scheduling points (the points on time line at which the scheduler makes decisions regarding which task is to be run next) are defined. According to this classification scheme, the three main types of schedulers are: clock-driven, event-driven, and hybrid. Important examples of event-driven schedulers are Earliest Deadline First (EDF) and Rate Monotonic analysis (RM). Event-driven schedulers are more sophisticated and usually more proficient and flexible than clock-driven schedulers. These are more proficient because they can feasibly schedule some task sets which clock-driven schedulers can not. These are more flexible because they can feasibly schedule sporadic and aperiodic tasks in addition to periodic tasks, whereas clock-driven schedulers can satisfactorily handle only periodic tasks\cite{mall}. Event-driven scheduling of real-time tasks was a subject of intense research during early 1970s, leading to the publication of a large number of research results, of which the following two popular algorithms are the essence of all EDF and RM\cite{w-s-liu}.

%Several variations to these two basic algorithms exist.

%We  focus  on  a  preemptive hard real-time systems in which periodic real-time tasks are scheduled with the Earliest-Deadline-First (EDF) algorithm or the Rate-Monotonic (RM) algorithm, the two most widely used real-time system models\cite{w-s-liu}.

\subsection{Earliest Dealine First (EDF)}
Earliest deadline scheduling is simple in concept. Every process tells the operating system scheduler its absolute time deadline. The scheduling algorithm simply allows the process that is in the greatest danger of missing its deadline to run first. Generally, this means that one process will run to completion if it has an earlier deadline than another. The only time a process would be preempted would be when a new process with an even shorter deadline becomes ready to run. To determine whether all the scheduled processes are capable of having their deadlines met, the following condition must hold:

$$\sum_{i = 1}^{n} \frac{C_i}{P_1} \leq 1$$

This simply tells us sum of all the percentages of CPU time used per process has to be less than or equal to 100%.

\subsection{Rate Monotonic (RM)}
 Rate monotonic analysis is a technique for assigning static priorities to periodic processes. As such, it is not a scheduler but a mechanism for governing the behavior of a preemptive priority scheduler. A conventional priority scheduler is used with this system, where the highest priority ready process will always get scheduled, preempting any lower priority processes.

A scheduler that is aware of rate monotonic scheduling would be provided with process timing parameters (period of execution) when the process is created and compute a suitable priority for the process. Most schedulers that support priority scheduling (e.g., Windows, Linux, Solaris, FreeBSD, NetBSD) do not perform rate monotonic analysis but only allow fixed priorities, so it is up to the user to assign proper priority levels for all real-time processes on the system. To do this properly, the user must be aware of all the real-time processes that will be running at any given time and each process' frequency of execution (1/T, where T is the period). To determine whether all scheduled processes can have their real-time demands met, the system has to also know each process' compute needs per period (C) and check that the following condition holds:

$$\sum_{i = 1}^{n} \frac{C_i}{P_1} < \ln 2$$

\section{Soft Real-Time Scheduling}

\section{Results}

\section{Related Work}

@@

\bibitem{chandrakasan}A.~Chandrakasan, S.~Sheng, and R.W.~Brodesen. \emph{Low-power CMOS digital design}. IEEE Journal of Solid-State Circuits, 27:473-484, April. 1992.

@@

\bibitem{moren}D.Moren. \emph{The next big thing in tech ought to be better batteries}. Macworld. August 19, 2016. Available at \url{http://www.macworld.com/article/3109692/hardware/the-next-big-thing-in-tech-ought-to-be-better-batteries.html}

@@

\bibitem{w-s-liu}W.-S.~Liu. \emph{Real-Time Systems}. Prentice Hall, Englewood Cliffs, NJ, June 2000.

@@

\bibitem{mall}R.~Mall. \emph{Real-Time Systems: Theory and practice}. Pearson Education India, 2009.

\bibitem{quian-andresen}H.~Qian and D.~Andresen, \emph{An energy-saving task scheduler for mobile devices}. 2015 IEEE/ACIS 14th International Conference on Computer and Information Science (ICIS), Las Vegas, NV, 2015, pp. 423-430.
doi: 10.1109/ICIS.2015.7166631

\bibitem{rao-et-al}V.~Rao, N.~Navet and G.~Singhal, \emph{Battery aware dynamic scheduling for periodic task graphs}. 2006 Proceedings 20th IEEE International Parallel \& Distributed Processing Symposium, Rhodes Island, 2006. doi: 10.1109/IPDPS.2006.1639403

\bibitem{takada-sakamura}H.~Takada and K.~Sakamura, \emph{Real-time synchronization protocols with abortable critical sections}. 1994 Proceedings of 1st International Workshop on Real-time Computing Systems \& Application, pp. 48-52

@@

\bibitem{weiser-et-al}M.~Weiser, B.~Welch, A.~Demers, and S.~Shenker. \emph{Scheduling for reduced CPU energy}. In Proc. of Symposium on Operating Systems Design and Implementation. Nov. 1994.

\bibitem{wilke-et-al}C.~Wilke, S.~Richly, S.~Götz, C.~Piechnick and U.~Aßmann, \emph{Energy Consumption and Efficiency in Mobile Applications: A User Feedback Study}, 2013 IEEE International Conference on Green Computing and Communications and IEEE Internet of Things and IEEE Cyber, Physical and Social Computing, Beijing, 2013, pp. 134-141.
doi: 10.1109/GreenCom-iThings-CPSCom.2013.45
